---
title: "R Notebook"
output: html_notebook
---

Initial look at data visualisation suggests models will not perform particuarly well at this stage.

However, there are plenty of things that could be tried with more promising data.

For example:

* Regression model to look at relative influence of different factors on total spend? 
* Unsupervised clustering algorithm to detect "groups" of customer profile that may spend differently?



This notebook has a very first look at kmeans clustering for customer segmentation.

Customer Segmentation is the subdivision of a market into discrete customer groups that share similar characteristics. Customer Segmentation can be a powerful means to identify unsatisfied customer needs. Using the above data companies can then outperform the competition by developing uniquely appealing products and services.

Ways in which businesses segment their customer base include:
1. Demographic information, such as gender, age, familial and marital status, income, education, and occupation.
2. Geographical information, which differs depending on the scope of the company. For localized businesses, this info might pertain to specific towns or counties. For larger companies, it might mean a customer’s city, state, or even country of residence.
3. Psychographics, such as social class, lifestyle, and personality traits.
4. Behavioral data, such as spending and consumption habits, product/service usage, and desired benefits.

Advantages of Customer Segmentation
*Determine appropriate product pricing.
*Develop customized marketing campaigns.
*Design an optimal distribution strategy.
*Choose specific product features for deployment.
*Prioritize new product development efforts.


Install following packages if you don't have them
```{r}
library("tidyverse")
library("scatterplot3d")
```


Read in data
```{r}
clarks <- read.csv("data/Data\ Science\ Assessment.csv")
```
```{r}
clarks <-  mutate(clarks,
                  total.spend = online.spend + store.spend) 

clarks$email <- as.numeric(clarks$email=="yes")
clarks[,3:14] <- scale(clarks[,3:14]) 
```



K-means algorithm can be summarized as follows:

Specify the number of clusters4ters (K) to be created (by the analyst)
Select randomly k objects from the data set as the initial clusters4ter centers or means
Assigns each observation to their closest centroid, based on the Euclidean distance between the object and the centroid
For each of the k clusters4ters update the clusters4ter centroid by calculating the new mean values of all the data points in the clusters4ter. The centroid of a Kth clusters4ter is a vector of length p containing the means of all variables for the observations in the kth clusters4ter; p is the number of variables.
Iteratively minimize the total within sum of square (Eq. 7). That is, iterate steps 3 and 4 until the clusters4ter assignments stop changing or the maximum number of iterations is reached. By default, the R software uses 10 as the default value for the maximum number of iterations.



3 clusters
```{r}
set.seed(20)
clusters3 <- kmeans(clarks[,c(3:6)], 3)

clarks$clusters3 <- as.factor(clusters3$cluster)

str(clusters3)
```


```{r}
 ggplot(data = clarks, aes(x = age, y = total.spend, colour = clusters3)) +
          geom_point(alpha = 0.1)+ guides(colour = guide_legend(override.aes = list(alpha = 1))) 
```



4 clusters
```{r}
set.seed(20)
clusters4 <- kmeans(clarks[,c(3:6)], 4)

clarks$clusters4 <- as.factor(clusters4$cluster)

str(clusters4)
```

5 clusters
```{r}
set.seed(20)
clusters5 <- kmeans(clarks[,c(3:6)], 5)

clarks$clusters5 <- as.factor(clusters5$cluster)

str(clusters5)
```


Note: there are established methods for choosing optimum number of clusters - most prominently Elbow, Silhouhette and Gap statistic methods.
Note: here I have used all data - would want to use a taining subset in reality.



Quick example of output:


Clustering was done on age, credit score, distance and email (age and credit score are loosely correlates so we will just plot age)
```{r}
scatterplot3d( clarks[,c(3,5,6)], color = as.numeric(clarks$clusters5))
```



If we remove email and instrad plot total spend we can see if our clusters have any relationship to total.spend - perhaps a little?
```{r}
scatterplot3d( clarks[,c(3,6,14)], color = as.numeric(clarks$clusters5))
```


We could explore other clustering algorithms. 
We could also look at regression or support vector machines to try to determine which factors effect total.spend and profile customers in this way.
We could also try using data we collect to predict if customers will be online only, store only or both which could feed into other recomendations


